---
title: "Project Tidy Data - Codebook template"
author: "Ray Lugo, Jr."
date: Sys.Date()
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Description
Short description of the project

##Study design and data processing

###Collection of the raw data
Description of how the data was collected.

###Notes on the original (raw) data 
Some additional notes (if avaialble, otherwise you can leave this section out).

##Creating the tidy datafile

###Guide to create the tidy data file
Description on how to create the tidy data file (1. download the data, ...)/

###Cleaning of the data
Short, high-level description of what the cleaning script does. [link to the readme document that describes the code in greater detail]()

##Description of the variables in the tiny_data.txt file
General description of the file including:
 - Dimensions of the dataset
 - Summary of the data
 - Variables present in the dataset
 (you can easily use Rcode for this, just load the dataset and provide the information directly form the tidy data file)

```{r cars}
summary(cars)
```

###Variable 1 (repeat this section for all variables in the dataset)
Short description of what the variable describes.

Some information on the variable including:
 - Class of the variable
 - Unique values/levels of the variable
 - Unit of measurement (if no unit of measurement list this as well)
 - In case names follow some schema, describe how entries were constructed (for example time-body-gyroscope-z has 4 levels of descriptors. Describe these 4 levels). 

(you can easily use Rcode for this, just load the dataset and provide the information directly form the tidy data file)

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

####Notes on variable 1:
If available, some additional notes on the variable not covered elsewehere. If no notes are present leave this section out.

##Sources
Sources you used if any, otherise leave out.

<!--chapter:end:codebook.Rmd-->

---
title: "Project Tidy Data - README.Rmd template"
author: "Ray Lugo, Jr."
date: Sys.Date()
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

<!--chapter:end:README.Rmd-->

---
title: "Project Tidy Data - run_analysis.R template"
author: "Ray Lugo, Jr."
date: Sys.Date()
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

<!--chapter:end:run_analysis.Rmd-->

