---
title: "Project Tidy Data - Codebook"
author: "Ray Lugo, Jr."
date: "November 25, 2016"
output:
  md_document:
    toc: yes
    toc_depth: 5
    variant: markdown_github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = '')
```
### Codebook Description

This codebook provides details on:

- Overview of the original dataset.
- Data processing method.
- Overview of run\_analysis.R
- Link to step-by-step commented documentation of run\_analysis.R
- Overview of the initial Tidy data set.
- Overview of the Summarized Tidy data set
- Data Dictionary of all variables in the summarized Tidy dataset.
- Initial statistics about the summarized tidy data set.

### Overview of the original dataset.

### Data Processing Method

### Overview of run\_analysis.R

Short, high-level description of what the cleaning script does. 

### Link to step-by-step commented documentation of run\_analysis.R

[link to the readme document that describes the code in greater detail]()

### Initial Tidy data set.


### Summarized Tidy data set

`r paste0(code.book.dtTidyAvg.tableName, " - ", code.book.dtTidyAvg.description, "\n\n")
`r paste0(code.book.dtTidyAvd.wording, "\n\n")


#### Listing of Variables

```{r, echo=FALSE}
eval(code.book.dtTidyAvg.columnDefs)
```

#### Detailed Variable Information

```{r, echo=FALSE}
eval(code.book.dtTidyAvg)
```

#### Initial statistics about the summarized tidy data set.


```{r pressure, echo=TRUE}
glimpse(dtTidyAvg)
```


#### Sources

https://gist.github.com/stevenworthington/3178163 <- function to load specified data files into data.tables withe the same name as the originating file.


