---
title: "Project Tidy Data - Codebook"
author: "Ray Lugo, Jr."
date: Sys.Date()
output:
  html_document:
    keep_md: yes
---
## Project Tidy Data Codebook

### Description

This codebook provides details on:

- Overview of the original dataset.
- Data processing method.
- Overview of run\_analysis.R
- Link to step-by-step commented documentation of run\_analysis.R
- Overview of the initial Tidy data set.
- Overview of the Summarized Tidy data set
- Data Dictionary of all variables in the summarized Tidy dataset.
- Initial statistics about the summarized tidy data set.

###Overview of the original dataset.

### Data Processing Method

### Overview of run\_analysis.R

Short, high-level description of what the cleaning script does. 

### Link to step-by-step commented documentation of run\_analysis.R

[link to the readme document that describes the code in greater detail]()

### Overview of the initial Tidy data set.

### Overview of the Summarized Tidy data set

### Data Dictionary of all variables in the Summarized Tidy dataset.


##Description of the variables in the tiny_data.txt file
General description of the file including:
 - Dimensions of the dataset
 - Summary of the data
 - Variables present in the dataset
 (you can easily use Rcode for this, just load the dataset and provide the information directly form the tidy data file)

```{r}
cat(code.book.dtTidyAvg.tableName, " - ", code.book.dtTidyAvg.description, "\n\n", sep = "")
cat(code.book.dtTidyAvd.wording, "\n\n", sep = "")
cat("----------\n\n", sep = "")
cat("Overview of Variables\n","________________\n", sep = "")
print(code.book.dtTidyAvg.columnDefs)
cat("\n\n\n")
print(code.book.dtTidyAvg)
```

### Initial statistics about the summarized tidy data set.

```{r pressure, echo=FALSE}
plot(pressure)
```


##Sources
Sources you used if any, otherise leave out.
